Reproducible Research: Peer Assessment 1
===============================================


## Loading and preprocessing the data
Load csv data 'activity.csv' and transform  the date data into a valid R date class . Assumes that the unziped data file 'activity.csv' is pesent in working dir.

```{r}

# Read in data file
actData <- read.csv("activity.csv") 

## In CSV Dates are in YYYY-MM-DD format, turn into Calendar date
actData$date <-as.Date(actData$date,"%Y-%m-%d")

```


First few records of loaded data

```{r}

head(actData)

```

Summary  of loaded data

```{r}

summary(actData)

```

## What is mean total number of steps taken per day?

Total number of steps per day 

```{r}

totalStepsPerDay <- tapply(actData$steps, actData$date,sum, na.rm=TRUE)

```

Histogram of 'total number of steps per day'

```{r}

hist(totalStepsPerDay,xlab="Total steps per Day", 
      ylab="frequency", main="Histogram of total steps per day")

```

The mean of total steps per day

```{r}

mean(totalStepsPerDay,na.rm=TRUE)

```


The median  of total steps per day

```{r}

median(totalStepsPerDay,na.rm=TRUE)

```


## What is the average daily activity pattern?
Time series on each interval averaged across all days( 61 days)

```{r}

#step per each interval averaged over all days
avgStepsPerInterval <- tapply(actData$steps,actData$interval,                            
                              mean,na.rm=TRUE)
uniqueIntervals <- row.names(avgStepsPerInterval)
plot(uniqueIntervals, avgStepsPerInterval, type="l",
     main="Average number of steps per interval across all days", 
     xlab="Intervals( of 5-mins)", ylab="Average # of steps across all days", 
     lwd=2 , col="64")

```


Time interval that contains maximum average number of steps over all days

```{r}

names(which.max(avgStepsPerInterval))

```

## Imputing missing values
Number of NA values in the  dataset

```{r}

sum(is.na(actData))

```

Impute   missing values with  the 'average  steps in  that interval across all days' and create a new data set.

```{r}

actDataNew <- actData;
for (  i in 1: nrow(actDataNew))
    {
       rw <- actDataNew[i,]
      
       if ( is.na(actDataNew[i,"steps"]))
            {
               actDataNew[i, "steps"] <- avgStepsPerInterval[paste("",actDataNew[i, "interval"], sep="")]
            }
    }

```

 First few rows of new Data set 
```{r}

head ( actDataNew)

```

Histogram on new dataset after Imputing missing values with Avg value.

```{r}

totalStepsPerDay <- tapply(actDataNew$steps, actDataNew$date,sum, na.rm=TRUE)
hist(totalStepsPerDay,xlab="Total steps per Day", 
      ylab="frequency", main="Histogram of total steps per day on imputed Data set")

```

The mean of total steps per day of Imputed dataset

```{r}

mean(totalStepsPerDay,na.rm=TRUE)

```


The median  of total steps per day of Imputed dataset

```{r}

median(totalStepsPerDay,na.rm=TRUE)

```


## Are there differences in activity patterns between weekdays and weekends?

```{r}

days <- weekdays(actDataNew$date)
actDataNew$dayType <- ifelse(days == "Saturday" | days == "Sunday", 
                                "Weekend", "Weekday")

avgStepsPerIntrPerDayType <- with(actDataNew, aggregate(x=steps, by=list(interval, dayType), 
                                                  FUN=function(x) mean(x, na.rm=T)))
names(avgStepsPerIntrPerDayType) <- c("interval","dayType","steps")
library(lattice)
xyplot(steps~interval | dayType, avgStepsPerIntrPerDayType,type="l",
       layout=c(1,2),xlab="Interval",ylab = "Avg  Numver of steps")


```

